---
title: "ETC5513 Assignment 1"
author: "Alex Delamer"
date: 2026/04/10
format: 
    html:
        theme:
            light: darkly
            dark: flatly
    pdf: 
        theme: default
number-sections: true
toc: true
embed-resources: true
lightbox: true
execute:
    echo: false
    output: false
    message: false
    warning: false
---

:::{.content-hidden}
All of the graphs in this report have a gray background so they aren't particularly jarring in either light or dark mode when rendered to HTML. This also means they don't look particularly nice in either mode either.

I tried to find a way to display a different plot depending on whether you're in light or dark mode; while this is pretty easy to do by just applying the .light-content and .dark-content classes to two code chunks, doing it this way completely breaks figure numbering and cross-referencing depending on which mode you're in. 

Any additional methods I might have tried seemed to involve messing directly with Javascript. See:

- https://github.com/orgs/quarto-dev/discussions/5439
- https://mickael.canouil.fr/posts/2023-05-30-quarto-light-dark/

This is scary and I didn't want to do it, but I still wanted to have both dark and light mode, so the graphs are gray.
:::


::: {.callout-tip .dark-content .content-hidden unless-format="html"}
This assignment also comes in light mode (when rendered to HTML)! ↗️

For information on why this tip is still visible when rendered to PDF, see [this forum discussion](https://github.com/quarto-dev/quarto-cli/issues/2140).
::: 

```{r setup}
library(tidyverse)
library(kableExtra)
library(knitr)
```

# Introduction

Having spent the last quarter running a personalised marketing campaign, we in the marketing department are now looking to justify that spending. Using marketing spend per customer and their subsequent purchase amount, we will be able to determine if a higher amount spent on advertising to a unique customer leads to that customer spending more money with our business. If we can find a positive correlation between marketing spend and purchase amount on a per customer basis, we should consider a continuation of this personalised marketing program in future as the most effective way of parting our customers from their money. 

Given that we service customers from the consumer to enterprise level, we will also examine if spending pattern response to personalised marketing spend changes by market segment, in the hope of finding the most effective ways to part our largest customers from their largest money.

# Research Question

Our mandated research question can be put simply:

*Did higher marketing spend per customer lead to higher purchase amounts?*

The secondary question we will also be looking to answer is:

*Did customer response to higher marketing spend vary across market segments?*

# The Data

## Dataset Introduction

```{r import and cleaning}
customer_data_raw <- read_csv("data/36749346_Alex Delamer.csv")

customer_data_clean <- customer_data_raw |> 
    select(marketing_spend, purchase_amount, segment) |>
    filter_out(is.na(purchase_amount) | is.na(marketing_spend))
```

We have collected the following data on our customers from the last quarter:

1. The customer ID we assigned to them
1. Their original signup date to our site
1. The marketing spend on this customer during the last quarter in AUD
1. The number of visits they made to the site last quarter
1. The amount the spent on our site after the last quarter in AUD
1. What market segment the customer belongs to

We will be using only items 3, 5 and 6 in our analysis to look for correlations between marketing spend and purchase amounts per customer across market segments during the last quarter.

## Dataset Description

Our cleaned dataset has the following characteristics:

1. It has `r ncol(customer_data_clean)` variables.
1. These variables are: *`r names(customer_data_clean)`*
1. There are `r nrow(customer_data_clean)` observations (rows) in the dataset.
    - Each observation represents a single customer
    - There are `r nrow(customer_data_raw) - nrow(customer_data_clean)` less observations than in the raw dataset; we have removed any customers that we were missing marketing spend or purchase amount values for.


Using our cleaned dataset, we can take our first look at any differences in market segments. As seen in @fig-purchase-spend-box-1 below, the average purchase size for each segment doesn't see much variance. Our biggest spenders actually come from the small-medium business sector rather than enterprise as might be expected. 

```{r creating boxplots}
customer_data_clean$segment <- factor(customer_data_clean$segment, levels = c("Consumer", "SMB", "Enterprise"))

purchasing_by_segment <- ggplot(customer_data_clean, aes(x = segment, y = purchase_amount, fill = segment)) +
    geom_boxplot(box.linewidth = 1, whisker.linewidth = 1) +
    ylim(0, 350) +
    theme(legend.position = "none", panel.background = element_blank(), axis.line = element_line(linewidth = 0.5, colour = "black"), plot.title = element_text(size = 20, hjust = 0.5), plot.background = element_rect(fill = "gray75"), panel.grid = element_blank()) +
    scale_fill_brewer(palette = "Dark2") +
    labs(x = "Market Segment", y = "Purchase Amount (AUD)")

spend_by_segment <- ggplot(customer_data_clean, aes(x = segment, y = marketing_spend, fill = segment)) +
    geom_boxplot(box.linewidth = 1, whisker.linewidth = 1) +
    ylim(0, 350) +
    theme(legend.position = "none", panel.background = element_blank(), axis.line = element_line(linewidth = 0.5, colour = "black"), plot.title = element_text(size = 20, hjust = 0.5), plot.background = element_rect(fill = "gray75"), panel.grid = element_blank()) +
    scale_fill_brewer(palette = "Dark2") +
    labs(x = "Market Segment", y = "Marketing Spend (AUD)")

total_marketing_spend <- as.character(sum(customer_data_clean$marketing_spend))
total_purchase_amount <- as.character(sum(customer_data_clean$purchase_amount))
```

```{r drawing boxplots}
#| output: true
#| label: fig-purchase-spend-box
#| fig-cap: "Purchase Amounts and Marketing Spend in AUD by Market Segment"
#| layout-ncol: 2
#| fig-subcap: 
#|   - "Purchases"
#|   - "Marketing Spend"

plot(purchasing_by_segment)
plot(spend_by_segment)
```

Enterprise also has the lowest mean purchase amount, as seen in @tbl-mean-purchases-by-segment. In what seems like some devastating news for both the marketing team and the business as a whole, we have a higher mean marketing spend than mean purchase amount in every market segment! 

To put it plainly: last quarter we spent $`r total_marketing_spend` in total on our personalised marketing campaign. The company made $`r total_purchase_amount` in revenue.

Whoops.

```{r #tbl-mean-purchases-by-segment}
#| output: true
#| tbl-cap: Mean purchase amount and market spend per market segment

mean_segment_purchases <- customer_data_clean |>
    mutate(Segment = segment) |>
    group_by(Segment) |> 
    summarise(`Number of Customers` = n(), `Mean Purchase (AUD)` = round(mean(purchase_amount), 2), `Mean Marketing Spend (AUD)` = round(mean(marketing_spend), 2))

kable(mean_segment_purchases) |>
    kable_styling(bootstrap_options = c("hover", "bordered"))
```

# Results

While we have just learned that the business is doomed due to marketing spending more to acquire every customer than we make from them (our bad on that one), we should at least still answer our initial question: **did all this marketing spending at least encourage our customers to buy more?**

In a word: *yes!!*

```{r #fig-spend-vs-purchases}
#| output: true
#| fig-cap: Marketing Spend vs Purchase amount per customer, separated by market segment

ggplot(customer_data_clean, aes(x = marketing_spend, y = purchase_amount, colour = segment)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    theme(legend.position = "none", panel.background = element_blank(), axis.line = element_line(linewidth = 0.5, colour = "black"), plot.title = element_text(size = 20, hjust = 0.5), plot.background = element_rect(fill = "gray75"), panel.grid = element_blank()) +
    facet_wrap(vars(segment), axes = "all") +
    labs(x = "Market Spend (AUD)", y = "Purchase Amount (AUD)") +
    scale_color_brewer(palette = "Dark2")
```

As seen in @fig-spend-vs-purchases, when we spend more money marketing to a customer the amount of money they spend with us tends to increase as well. Assuming that turning a profit is not a priority (or indeed, something we need to do at all), continuing to spend larger and larger amounts on personalised marketing should see us continue to pump our gross revenue numbers higher and higher!

We have also been able to answer our secondary question; this trend seems to hold across all three of our market segments. 

## Caveats

```{r linear models}
overall_model <- lm(purchase_amount ~ marketing_spend, data = customer_data_clean)
consumer_model <- lm(purchase_amount ~ marketing_spend, data = customer_data_clean, subset = (segment == "Consumer"))
smb_model <- lm(purchase_amount ~ marketing_spend, data = customer_data_clean, subset = (segment == "SMB"))
enterprise_model <- lm(purchase_amount ~ marketing_spend, data = customer_data_clean, subset = (segment == "Enterprise"))
```

As we can see in @tbl-r-squared-values, the linear models that are represented by the fitted lines in @fig-spend-vs-purchases are good but not perfect predictors of customer purchasing behaviour. When all segments are considered together, marketing spend accounts for `r round(summary(overall_model)$r.squared, 4) * 100`% of the variance in customer purchasing habits. However, in the consumer segment where we have the highest number of customers, our model only accounts for `r round(summary(consumer_model)$r.squared, 4) * 100`% of variance. 

```{r tbl-r-squared-values}
#| output: true
#| tbl-cap: R-Squared values for linear models of purchase amount with marketing spend as predictor

r_squared_summary_table <- data.frame("Market Segment" = c("All Segments", "Consumer", "SMB", "Enterprise"), "R-Squared Value" = c(round(summary(overall_model)$r.squared, 4), round(summary(consumer_model)$r.squared, 4), round(summary(smb_model)$r.squared, 4), round(summary(overall_model)$r.squared, 4)), check.names = FALSE)

kable(r_squared_summary_table) |> 
    kable_styling(bootstrap_options = c("hover", "bordered"))
```

It seems possible that this correlation between marketing spend and purchase amount is an artifact of additional factors not recorded in the dataset. Consider the following:

- Did we spend more on personalised advertising to customers who had previously spent a lot of money with us and were likely to do so again even without the increased marketing push?
- Were these purchases even affected by marketing spend? We didn't record if a customer made their purchases this quarter before or after we spent the money to advertise to them.
- Did the type of personalised advertising a customer received affect the size of their purchase? We used ads, emails and coupons but didn't record which of those and in what quantities a given customer received.